\racets: Faceted Execution in Racket

### 2018-07-24 paper/paper.tex
colors for ell

Although \code{mark-hit} will operate on sensitive data (the game
boards), it is written without any special machinery to maintain the
secrecy of \code{board}. Protecting data {w.r.t.} policies is instead handled
automatically and implicitly by a runtime monitor.
When Alice and Bob want to play a game, they both create a
\emph{label} to protect their data. A label is unique id mapped to a policy predicate
that takes a key (e.g., the current user's name) and returns true or false to indicate
permission for the label. Alice's label is used to annotate the data she wants
to be kept secret. Supposing Alice chooses to be player 1, she may use the
following label:

@@

This code illustrates label creation, policy predicates, and the first-class
nature of labels. The policy predicate $\texttt{($\lambda$ (\blue{x}) (= 1 \blue{x}))}$
grants permission to player 1 only and is associated with the dynamically generated
label l (returned and bound to \texttt{\blue{alice-label}}).
Bob would use a similar policy (but for player $2$ instead of $1$). At runtime,
the \code{let-label} form creates a label \blue{$\ell_{A}$} and binds
it to a closure for its policy predicate. When Alice wants to protect a
value, she creates a facet annotated with her label and two
\emph{branches}. The positive (left) branch represents the value as it
should appear to her, and the negative (right) to everyone else:

@@

In the above example, \blue{$\star$} (lazy failure) is used in the negative branch
to ensure execution will fail if Bob tries to observe Alice's secret gameboard. To observe
Alice's gameboard, Bob can try to use $\code{(obs}\ \blue{e_\ell}\ \blue{e_\text{key}}\ 
\blue{e_\text{fac}}\code{)}$ form, which takes a label, a key, and a faceted value.
Explicit observation projects a single label $\blue{e_\ell}$ in faceted value $\blue{e_\text{fac}}$ 
to either its positive or negative facet, depending on whether the policy associated 
with $\blue{e_\ell}$ returns true for key $\blue{e_\text{key}}$. If Bob tries to observe 
Alice's board, the policy predicate will return false (from \texttt{(= 1 $\blue{2}$)}) 
and Alice's negative facet $\blue{\star}$ will result.

In other applications, Alice may choose a sensible default value to
reveal to others---she may even want to create a \emph{nested}
facet. For example, a social-networking application may use a
nested facet consisting of two labels for
\blue{$\ell_{\text{Friends}}$} and \blue{$\ell_{\text{Family}}$}. A user
can then present three views of her social-media profile: \bmth{p_1} to
her family, containing her phone number and other contact information,
\bmth{p_2} to her friends showing her interests, and \bmth{p_3} to
everyone else, showing only her name and email.

@@

Because the applied function can be stateful, faceted execution also
maintains the current privilege level in a \emph{program counter} (PC). The
program counter is a property of the current evaluation context and is used
to build facets when writes are made to the store in a privileged context.
For example, if a stateful function ``splits'' when applied on both the positive and negative
facets of a value faceted by a label $\blue{\ell}$, and on the positive branch
the function uses $\texttt{set!}$ to mutate a variable $x$ from $2$ to $3$,
FE semantics will set $x$ to $\facet{\ell}{3}{2}$ so that the value $3$
cannot be leaked from the secure context (speculative execution under the
$\blue{+\ell}$ facet). This is because, for the duration of the app ``split'',
the evaluation context records that all values are implicitly guarded by $\blue{+\ell}$
and then $\blue{-\ell}$, respectively. If the semantics for $\texttt{set!}$ does
not make this faceting explicit, a sensitive value can leak from one PC to another.
We expand upon these subtleties in Section~\ref{sec:semantics},
where we present a complete semantics for faceted execution.

@@

@@

Values in our semantics are either (unfaceted) base values or facets
composed of a label and two branches. Facets can nest, allowing trees
of faceted values. We use the term \emph{branches} to refer to
positive or negated labels. Collections of branches define the program
counter \rmth{pc}, which tracks the set of branches in the current
context. For example, to apply a faceted function to a value (as in
the application of \bmth{\facet{\ell}{\lambda x.~0}{\lambda x.~1}}),
the semantics first applies \bmth{\lambda x.~0} while extending \rmth{pc}
with \rmth{+\ell}, then applies the negative branch extending
\rmth{pc} with \rmth{-\ell}.

@@

Similarly, program counters are sets of branches. However, we must
still ask how we will keep track of the ``current'' program
counter. Our implementation uses Racket's parameters, though other
mechanisms (such as continuation marks\kris{cite}, to which parameters
macro-expand) can also be used. \racets defines the parameter
\code{current-pc}, and updates it as computation progresses:

@@

There is a wealth of existing Racket code we may like to incorporate
into \racets programs. For example, our case study uses the
\texttt{web-server} framework for building web applications. However,
in general, we believe that interacting with code not written using
faceted execution is a challenging open problem, and we do not know of any
principled solutions in the literature.

@@

Several other efforts into dynamic analysis for information flow are
worth noting. \citet{Stefan:2011} first presented \code{LIO}---a monad
(with implementation in Haskell) that tracks privilege of the current
program counter and forbids effects that would violate the security
policy. It may be surprising that \code{LIO}
works well for Haskell programs, given that faceted execution is
more precise than \code{LIO}---allowing values to become faceted
rather than halting the program. One key difference is that Haskell
programs emphasize purity while languages such as
JavaScript (the original target of faceted execution) does not, so much
of the machinery for faceted execution's effect on the store is less
interesting. Several authors have implemented related systems to
\code{LIO}, including variants of faceted execution \cite{Schmitz:2016}
and variants of \code{LIO} that extend its
power to arbitrary monad transformers~\cite{james}. We believe that it
would be possible to implement a variant of our technique that would
give similar insights to programs using \code{LIO}, though much of the
interesting machinery for handling state may be unnecessary.

>>>>

Although \code{mark-hit} will operate on sensitive data (the game
boards), it is written without any special machinery to maintain the
secrecy of \code{board}. Protecting data {w.r.t.} policies is instead handled
automatically and implicitly by a runtime monitor.
When Alice and Bob want to play a game, they both create a
\emph{label} to protect their data. A label is unique id mapped to a policy predicate
that takes an key (e.g., the current user's name) and returns true or false to indicate
permission for the label. Alice's label is used to annotate data she wants
to be kept secret. Supposing Alice chooses to be player 1, she may use the
following label:

@@

This code illustrates label creation, policy predicates, and the first-class
nature of labels. The policy predicate $\texttt{($\lambda$ (\blue{x}) (= 1 \blue{x}))}$
grants permission to player 1 only and is associated with the dynamically generated
label l (returned and bound to \texttt{\blue{alice-label}}).
Bob would use a similar policy (but for player $2$ instead of $1$). At runtime,
the \code{let-label} form creates a label \fcol{$\ell_{A}$} and binds
it to a closure for its policy predicate. When Alice wants to protect a
value, she creates a facet annotated with her label and two
\emph{branches}. The positive (left) branch represents the value as it
should appear to her, and the negative (right) to everyone else:

@@

In the above example, \blue{$\star$} (lazy failure) is used in the negative branch
to ensure execution will fail if Bob tries to observe Alice's secret gameboard.
The $\code{(obs}\ \fcol{e_\ell}\ \blue{e_\text{key}}\ \blue{e_\text{fac}}\code{)}$ form takes
a label, a key, and a faceted value. Explicit observation projects a single label
$\fcol{e_\ell}$ in faceted value $\blue{e_\text{fac}}$ to either its positive or negative facet,
depending on whether the policy associated with $\fcol{e_\ell}$ returns true for key
$\blue{e_\text{key}}$. If Bob tries to observe Alice's board, the policy predicate
will return false (from \texttt{(= 1 $\blue{2}$)}) and Alice's negative facet $\blue{\star}$ will result.

In other applications, Alice might choose a sensible default value to
reveal to others---she may even want to create a \emph{nested}
facet. For example, a social-networking application might use a
nested facet consisting of two labels for
\fcol{$\ell_{\text{Friends}}$} and \fcol{$\ell_{\text{Family}}$}. A user
could then present three views of her social-media profile: \bmth{p_1} to
her family, containing her phone number and other contact information,
\bmth{p_2} to her friends showing her interests, and \bmth{p_3} to
everyone else, showing only her name and email.

@@

Because the applied function can be stateful, faceted execution also
maintains the current privilege level in a \emph{program counter} (PC). The
program counter is a property of the current evaluation context and is used
to build facets when writes are made to the store in a privileged context.
For example, if a stateful function ``splits'' when applied on both the positive and negative
facets of a value faceted by a label $\fcol{\ell}$, and on the positive branch
the function uses $\texttt{set!}$ to mutate a variable $x$ from $2$ to $3$,
FE semantics will set $x$ to $\facet{\ell}{3}{2}$ so that the value $3$
cannot be leaked from the secure context (speculative execution under the
$\fcol{+\ell}$ facet). This is because, for the duration of the the app ``split'',
the evaluation context records that all values are implicitly guarded by $\fcol{+\ell}$
and then $\fcol{-\ell}$, respectively. If the semantics for $\texttt{set!}$ does
not make this faceting explicit, a sensitive value could leak from one PC to another.
We expand upon these subtleties in Section~\ref{sec:semantics},
where we present a complete semantics for faceted execution.

@@

If we had passed any value other than $1$ into the \code{obs} form,
Alice's policy would return \code{false}. In this case, the \code{obs}
form would reduce to $\blue{\star}$, and the program would fail
(indicating that someone other than Alice attempted to view her
board).

@@

Values in our semantics are either (unfaceted) base values or facets
composed of a label and two branches. Facets can nest, allowing trees
of faceted values. We use the term \emph{branches} to refer to
positive or negated labels. Collections of branches define the program
counter \rmth{pc}, which tracks the set of branches in the current
context. For example, to apply a faceted function to a value (as in
the application of $\facet{\ell}{\blue{\lambda x.~0}}{\blue{\lambda x.~1}}$),
the semantics first applies \bmth{\lambda x.~0} while extending \rmth{pc}
with \rmth{+\ell}, then applies the negative branch extending
\rmth{pc} with \rmth{-\ell}.

@@

Similarly, program counters are sets of branches. However, we must
still ask how we will keep track of the ``current'' program
counter. Our implementation uses Racket's parameters, though other
mechanisms (such as continuation marks\kris{cite}, to which parameters
macro-expand) could also be used. \racets defines the parameter
\code{current-pc}, and updates it as computation progresses:

@@

There is a wealth of existing Racket code we might like to incorporate
into \racets programs. For example, our case study uses the
\texttt{web-server} framework for building web applications. However,
in general, we believe that interacting with code not written using
faceted execution is a challenging open problem, and we do not know of any
principled solutions in the literature.

@@

Several other efforts into dynamic analysis for information flow are
worth noting. \citet{Stefan:2011} first presented \code{LIO}---a monad
(with implementation in Haskell) that tracks privilege of the current
program counter and forbids effects that would violate the security
policy. It might be surprising that \code{LIO}
works well for Haskell programs, given that faceted execution is
more precise than \code{LIO}---allowing values to become faceted
rather than halting the program. One key difference is that Haskell
programs emphasize purity while languages such as
JavaScript (the original target of faceted execution) does not, so much
of the machinery for faceted execution's effect on the store is less
interesting. Several authors have implemented related systems to
\code{LIO}, including variants of faceted execution \cite{Schmitz:2016}
and variants of \code{LIO} that extend its
power to arbitrary monad transformers~\cite{james}. We believe that it
would be possible to implement a variant of our technique that would
give similar insights to programs using \code{LIO}, though much of the
interesting machinery for handling state may be unnecessary.
